---
title: "R Notebook"
output: html_notebook
---

#--------------------------------  NOTE  ----------------------------------------
# 1 This code is to quantify single time series uncertainty based on different data sources;
# 2 The format of the input data is [ERA OBS];
# 3 Coder: Flora Zhang         Date: 08/02/2023       @ NREL
#--------------------------------------------------------------------------------
## packages double check

```{r}
temporal <- function(df) {
  library(ForeCA)
  library(fields)
  library(imputeTestbench)
  library(zoo)
  library(MLmetrics)
  library(zoo)
  library(ForeCA)
  library(fields)
  library(hydroGOF)
  # df <- data.frame(dataset[,c('ERA', 'OBS')])
  # window size: 12
  
  # Moving-window Entropy
  en_era <- c()
  en_obs <- c()
  for (i in 1:(dim(df)[1]-12+1)){
    en_era[i] <- ForeCA::spectral_entropy(df$ERA[i:(i+11)])
    en_obs[i] <- ForeCA::spectral_entropy(df$OBS[i:(i+11)])
  } 
  
  # Moving-window Standard Deviation
  sd_era <- c()
  sd_obs <- c()
  for (i in 1:(dim(df)[1]-12+1)){
    sd_era[i] <- sd(df$ERA[i:(i+11)])
    sd_obs[i] <- sd(df$OBS[i:(i+11)])
  } 
  
  # Moving-window Turbulence Intensity
  ti_era <- c()
  ti_obs <- c()
  for (i in 1:(dim(df)[1]-12+1)){
    ti_era[i] <- (max(df$ERA[i:(i+11)])-min(df$ERA[i:(i+11)]))/mean(df$ERA[i:(i+11)])
    ti_obs[i] <- (max(df$OBS[i:(i+11)])-min(df$OBS[i:(i+11)]))/mean(df$OBS[i:(i+11)])
  } 
  
  # Moving-window Variability Index
  vi_era <- c()
  vi_obs <- c()
  for (i in 1:(dim(df)[1]-12+1)){
    vi_era[i] <- sd(df$ERA[i:(i+11)])/mean(df$ERA[i:(i+11)])
    vi_obs[i] <- sd(df$OBS[i:(i+11)])/mean(df$OBS[i:(i+11)])
  } 
  
  #result <- list('entropy' = cbind(en_era, en_obs),  'sd' = cbind(sd_era, sd_obs), 
  #               'ti' = cbind(ti_era, ti_obs), 'vi' = cbind(vi_era, vi_obs))
  result <- as.matrix(cbind(en_era, en_obs, sd_era, sd_obs, ti_era, ti_obs,vi_era, vi_obs))
  colnames(result) <- c('entropy_era', 'entropy_obs','sd_era', 'sd_obs','ti_era','ti_obs','vi_era', 'vi_obs')
  return(result)
}
```

Here we load the Argonne example data. Note that:
1. The in situ measurement does include some missing values. So we linearly interpolate these missing values. 
2. We compute the metrics based on ERA5 and observations, separately. 
```{r}
argonne2018 <- read.csv("~/Desktop/sample_data/il_argonne_10m_2018.csv")
df <- data.frame(ERA = argonne2018$ERA510m, OBS = na.approx(argonne2018$Obs10m))
result <- data.frame(temporal(df))
result
```

Here we visualize the pairwise metrics based on ERA5 and observation in heatmaps. These heatmaps are by hour of the day and by day of the year so that we could see both the diurnal and seasonal patterns from these metrics.  
```{r}
en_era <- result$entropy_era
en_obs <- result$entropy_obs
sd_era <- result$sd_era
sd_obs <- result$sd_obs
ti_era <- result$ti_era
ti_obs <- result$ti_obs
vi_era <- result$vi_era
vi_obs <- result$vi_obs

en_era.matrix <- matrix(en_era, nrow = 365, ncol = 24)
sd_era.matrix <- matrix(sd_era, nrow = 365, ncol = 24)
ti_era.matrix <- matrix(ti_era, nrow = 365, ncol = 24)
vi_era.matrix <- matrix(vi_era, nrow = 365, ncol = 24)

en_obs.matrix <- matrix(en_obs, nrow = 365, ncol = 24)
sd_obs.matrix <- matrix(sd_obs, nrow = 365, ncol = 24)
ti_obs.matrix <- matrix(ti_obs, nrow = 365, ncol = 24)
vi_obs.matrix <- matrix(vi_obs, nrow = 365, ncol = 24)

set.panel(4,2)
image.plot(1:365, 1:24, en_era.matrix, xlab = "day", ylab = "hour", main = "Entropy")
image.plot(1:365, 1:24, en_obs.matrix, xlab = "day", ylab = "hour", main = "Entropy")

image.plot(1:365, 1:24, sd_era.matrix, xlab = "day", ylab = "hour", main = "Standard Deviation")
image.plot(1:365, 1:24, sd_obs.matrix, xlab = "day", ylab = "hour", main = "Standard Deviation")

image.plot(1:365, 1:24, ti_era.matrix, xlab = "day", ylab = "hour", main = "Turbulence Intensity")
image.plot(1:365, 1:24, ti_obs, xlab = "day", ylab = "hour", main = "Turbulence Intensity")

image.plot(1:365, 1:24, vi_era.matrix, xlab = "day", ylab = "hour", main = "Variability Index")
image.plot(1:365, 1:24, vi_obs.matrix, xlab = "day", ylab = "hour", main = "Variability Index")
```
Besides visualization, we also want to quantify the difference/similarity between metrics calculates based on ERA5 and observations. Here we adopt cosine similarity where the values closer to one indicates similarity and the values closer to zero indicates the two vectors are orthogonal. 
```{r}
library(lsa)
cosine(en_era, en_obs)
cosine(sd_era, sd_obs)
cosine(ti_era, ti_obs)
cosine(vi_era, vi_obs)
```